RunTime-Assisted Convergence in Replicated Data Types

### 2022-04-25 abstract-semantics.tex
Committing edits

\paragraph{Properties} We now formalize the notable properties of the
abstract machine and its executions.

@@

\begin{theorem}[{\bf Progress}]
  \label{thm:progress}
  Every reachable state $\Delta = ((V,E),\\N,C,H,L)$ of the abstract
  machine is either: 
  \begin{itemize}
    \item A ``quiescent'' state, where: $\forall b_1,b_2 \in
      dom(H).~C(H(b_1)) = C(H(b_2))$, Or
    \item An ``unstuck'' state, where there exist $b_1,b_2 \in dom(H)$
      that satisfy the pre-conditions of \rulelabel{Merge} or
      \rulelabel{FastFwd} rules, i.e., $b_1$ and $b_2$ are mergeable. 
  \end{itemize}
\end{theorem}

@@

\noindent\paragraph{Ivy formalization} Appendix contains the manual
proofs of the aforementioned theorems. In addition we also formalized
the \quark abstract machine in Ivy~\cite{ivy}, and verified its
properties automatically with help of Z3~\cite{z3}. Ivy is a
multi-modal verification tool, whose primary purpose is to verify the
design and implementations of distributed protocols. The key
differentiating aspect of Ivy is its onus on predictable automation
and decidable reasoning.  Automated proofs in Ivy are restricted to
logical fragments for which the tool is a decision procedure. %Towards
% this end, Ivy checks the verification conditions for syntactic
% confirmity with known decidable fragments, and provides feedback to
% the user if the checks fail. The feedback helps users reorganize their
% formal development or reformulate their queries until the reasoning
% becomes decidable. Ivy uses induction to prove the properties of
% infinite-state systems. The failure to discharge a proof results in a
% counterexample, which is either a real counterexample (i.e., a bug in
% the system) or a counterexample to induction. In the latter case, the
% counterexample serves as a guide for the user to formulate the
% inductive invariants required to discharge the proof. 
Consequently, the creative aspect of proving a theorem in Ivy lies in
discovering decidable abstractions that characterize the system being
reasoned about, and crafting the inductive invariants required to
discharge the automated proof. This is in contrast to theorem provers
such as Coq and Isabelle, and their extensions such as
Sledgehammer~\cite{sledge}, where automation is used on a best-effort
basis to aid interactive theorem proving.

>>>>

\paragraph{Properties} We now formalize the notable properties of the
abstract machine and its executions\footnote{Proofs can be found in
the Appendix}.

@@

\begin{theorem}[{\bf Progress}]
  \label{thm:progress}
  Every reachable state $\Delta = ((V,E),\\N,C,H,L)$ of the abstract
  machine is either: 
  \begin{itemize}
    \item A ``quiescent'' state, where: $\\\forall b_1,b_2 \in
      dom(H).~C(H(b_1)) = C(H(b_2))$, Or
    \item An ``unstuck'' state, where there exist $b_1,b_2 \in dom(H)$
      that satisfy the pre-conditions of \rulelabel{Merge} or
      \rulelabel{FastFwd} rules, i.e., $b_1$ and $b_2$ are mergeable. 
  \end{itemize}
\end{theorem}

@@

\noindent\paragraph{Ivy formalization}  In addition to the manual
proofs, we also formalized the \quark abstract machine in
Ivy~\cite{ivy} and verified its properties automatically with help of
Z3~\cite{z3}. Ivy is a multi-modal verification tool, whose primary
purpose is to verify the design and implementations of distributed
protocols. The key differentiating aspect of Ivy is its onus on
predictable automation and decidable reasoning.  Automated proofs in
Ivy are restricted to logical fragments for which the tool is a
decision procedure. %Towards
% this end, Ivy checks the verification conditions for syntactic
% confirmity with known decidable fragments, and provides feedback to
% the user if the checks fail. The feedback helps users reorganize their
% formal development or reformulate their queries until the reasoning
% becomes decidable. Ivy uses induction to prove the properties of
% infinite-state systems. The failure to discharge a proof results in a
% counterexample, which is either a real counterexample (i.e., a bug in
% the system) or a counterexample to induction. In the latter case, the
% counterexample serves as a guide for the user to formulate the
% inductive invariants required to discharge the proof. 
Consequently, the creative aspect of proving a theorem in Ivy lies in
discovering decidable abstractions that characterize the system being
reasoned about, and crafting the inductive invariants required to
discharge the automated proof. This is in contrast to theorem provers
such as Coq and Isabelle, and their extensions such as
Sledgehammer~\cite{sledge}, where automation is used on a best-effort
basis to aid interactive theorem proving.

### 2022-04-25 abstract.tex
Committing edits

We propose a runtime-assisted approach to enforce convergence in distributed executions of replicated data types. The key distinguishing aspect of our approach is that it guarantees convergence \emph{unconditionally} -- without requiring data type operations to satisfy algebraic laws such as commutativity and idempotence. Consequently, programmers are no longer obligated to prove convergence on a per-type basis. Moreover, our approach lets sequential data types be reused in a distributed setting by extending their implementations rather than refactoring them.  The novel component of our approach is a distributed runtime that orchestrates \emph{well-formed } executions that are guaranteed to converge. Despite the utilization of a runtime, our approach comes at no additional cost of latency and availability. We implement our approach in a system called \quark and conduct a thorough evaluation of its performance over a large-scale collaborative editing application.

>>>>

We propose a runtime-assisted approach to enforce convergence in distributed executions of replicated data types. The key distinguishing aspect of our approach is that it guarantees convergence \emph{unconditionally} -- without requiring data type operations to satisfy algebraic laws such as commutativity and idempotence. Consequently, programmers are no longer obligated to prove convergence on a per-type basis. Moreover, our approach lets sequential data types be reused in a distributed setting by extending their implementations rather than refactoring them.  The novel component of our approach is a distributed runtime that orchestrates \emph{well-formed } executions that are guaranteed to converge. Despite the utilization of a runtime, our approach comes at no additional cost of latency and availability. Instead, we introduce a novel tradeoff against a metric called \emph{staleness}, which roughly corresponds to the time taken for replicas to converge. We implement our approach in a system called Quark and conduct a thorough evaluation of its tradeoffs.

### 2022-04-25 concrete-semantics.tex
Committing edits

\paragraph{Refinement} The following theorem asserts that the \quark
distributed machine refines the abstract machine described in
Sec.~\ref{sec:abstract-sem}:
\begin{theorem}[{\bf Refinement}]
  \label{thm:refinement}
  There exists a refinement relation $R$ relating abstract and
  concrete states such that:
  \begin{itemize}
    \item $R$ relates initial abstract and concrete states, i.e.,
      $R(\Delta_{\odot},\delta_{\odot})$.
    \item Forall $\delta,\delta',\Delta$, if $R(\delta,\Delta)$ and
      $\delta \qstepsto \delta'$, then there exists a $\Delta'$ such
      that $\Delta \stepsto \Delta'$ and $R(\delta',\Delta')$.
  \end{itemize}
\end{theorem}
The proof is done in Ivy. The refinement relation $R$ essentially
relates the version vectors of concrete semantics to the version
histories of abstract semantics. A \emph{happens-before} relation has
been defined as a ghost state in concrete semantics to help us
establish the refinement relation in Ivy.

>>>>

\paragraph{Refinement} The relationship between the \quark abstract
(Sec.~\ref{sec:abstract-sem}) and distributed
(Fig.~\ref{fig:quark-semantics}) machines is stated thus:
\begin{theorem}[{\bf Refinement}]
  \label{thm:refinement}
  There exists a refinement relation $R$ relating abstract and
  concrete states such that:
  \begin{itemize}
    \item $R$ relates the initial abstract and concrete states, i.e.,
      $R(\Delta_{\odot},\delta_{\odot})$.
    \item Forall $\delta,\delta',\Delta$, if $R(\delta,\Delta)$ and
      $\delta \qstepsto \delta'$, then there exists a $\Delta'$ such
      that $\Delta \stepsto \Delta'$ and $R(\delta',\Delta')$.
  \end{itemize}
\end{theorem}
The proof is done in Ivy. The refinement relation $R$ essentially
relates the version vectors of concrete semantics to the version
histories of abstract semantics. A \emph{happens-before} relation has
been defined as a ghost state in concrete semantics to help us
establish the refinement relation in Ivy.

### 2022-04-25 evaluation.tex
Committing edits

\begin{figure}[ht]
  \centering
    \includegraphics[scale=0.4]{Figures/monkey_latency}
\caption{Latency of writes to a shared document under \quark. }
\label{fig:latency}
\vspace*{-0.2in}
\end{figure}

@@

\begin{figure}[ht]
  \centering
    \includegraphics[scale=0.4]{Figures/monkey_staleness}
\caption{Staleness increases as the number of concurrent editors
  increase.}
\label{fig:monkey-staleness}
% \vspace*{-0.2in}
\end{figure}

To understand the effect of \quark on
staleness, we quantify and measure it along with latency in the
experiment setup described above. Staleness is defined as the time
taken for a version committed on one replica to be merged into a
concurrent version on a remote replica. To measure staleness, we
annotate every version $v$ with the timestamp $t$ of the wall clock
time of its creation. When $v$ is is merged into a remote branch $b$
at a later time $t'$, the difference $t' - t$ denotes the staleness of
$v$ w.r.t the new version on $b$. One such staleness measurement is
recorded for every merge that ever happens during the experiment. We
compute 10th, 50th, and 90th percentiles of staleness values thus
obtained. Fig.~\ref{fig:monkey-staleness} shows the results. As
evident, staleness increases steadily with the increasing number of
concurrent replicas, which is expected considering that merges are
synchronized, and increasing the number of replicas reduces the number
of opportunities for a replica to merge. The increase is roughly
linear in the number of replicas as our implementation passes the lock
around in a round-robin fashion. Despite the steady increase, the 90th
percentile staleness values remain low -- in the order of 100ms with
the number of replicas under 30, and in the order of 1s with number of
replicas under 60. While further optimizations might reduce staleness,
a non-trivial staleness overhead is inevitable in \quark due to our
use of synchronized merges to guarantee convergence.

@@

\begin{figure}[ht]
  \centering
    \includegraphics[scale=0.4]{Figures/rbmonkey_staleness}
  \caption{Staleness of key-value store merge operations.}
\label{fig:rb-staleness}
% \vspace*{-0.2in}
\end{figure}

>>>>

\begin{figure}[ht]
  \centering
    \includegraphics[scale=0.4]{Figures/monkey_latency}
\caption{Latency of writes to a shared document under \quark. }
\label{fig:latency}
\vspace*{-0.1in}
\end{figure}

@@

\begin{figure}[ht]
  \centering
    \includegraphics[scale=0.4]{Figures/monkey_staleness}
\caption{Staleness increases as the number of concurrent editors
  increase.}
\label{fig:monkey-staleness}
\vspace*{-0.2in}
\end{figure}

To understand the effect of \quark on
staleness, we quantify and measure it along with latency in the
experiment setup described above. Staleness is defined as the time
taken for a version committed on one replica to be merged into a
concurrent version on a remote replica. To measure staleness, we
annotate every version $v$ with the timestamp $t$ of the wall clock
time of its creation. When $v$ is is merged into a remote branch $b$
at a later time $t'$, the difference $t' - t$ denotes the staleness of
$v$ w.r.t the new version on $b$. One such staleness measurement is
recorded for every merge that ever happens during the experiment. We
compute 10th, 50th, and 90th percentiles of staleness values thus
obtained. Fig.~\ref{fig:monkey-staleness} shows the results. As
evident, staleness increases steadily with the increasing number of
concurrent replicas, which is expected considering that merges are
synchronized, and increasing the number of replicas reduces the number
of opportunities for a replica to merge. The increase is roughly
linear in the number of replicas as our implementation passes the lock
around in a round-robin fashion. Despite the steady increase, the 90th
percentile staleness values remain low -- less than 1s with the
number of replicas under 30, and less than 2s for number of replicas
under 60. While further optimizations might reduce staleness, a
non-trivial staleness overhead is inevitable in \quark due to our use
of synchronized merges to guarantee convergence.

@@

\begin{figure}[ht]
  \centering
    \includegraphics[scale=0.4]{Figures/rbmonkey_staleness}
  \caption{Staleness of key-value store merge operations.}
\label{fig:rb-staleness}
%\vspace*{-0.2in}
\end{figure}

### 2022-04-25 quark.tex
Committing edits

%\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{ifpdf}
\usepackage{graphicx}
%\usepackage[usenames,dvipsnames]{color}
\usepackage{stmaryrd}
%\usepackage[numbers]{natbib}
\usepackage{amsthm}
\usepackage{listings}          % format code
\usepackage{wrapfig}
\usepackage{textcomp}
\usepackage{tabularx}
\usepackage{color}
\usepackage{url}
\usepackage{tikz}
\usepackage{multirow,array}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}

@@

%% Title information
\title[]{Runtime-Assisted Convergence in Replicated Data Types}
%% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

@@

%% Author with single affiliation.
\author{Gowtham Kaki}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position1}
%  \department{Department1}              %% \department is recommended
  \institution{University of Colorado Boulder}            %% \institution is required
%  \streetaddress{Street1 Address1}
%  \city{City1}
%  \state{State1}
%  \postcode{Post-Code1}
  \country{USA}                    %% \country is recommended
}
%\email{first1.last1@inst1.edu}          %% \email is recommended

%% Second Author
\author{Prasanth Prahladan}
%\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position2a}
%  \department{Department2a}             %% \department is recommended
  \institution{University of Colorado Boulder}           %% \institution is required
%  \streetaddress{Street2a Address2a}
%  \city{City2a}
%  \state{State2a}
%  \postcode{Post-Code2a}
  \country{USA}                   %% \country is recommended
}
%\email{first2.last2@inst2a.com}         %% \email is recommended

%% Third Author
\author{Nicholas Lewchenko}
%\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position2a}
%  \department{Department2a}             %% \department is recommended
  \institution{University of Colorado Boulder}           %% \institution is required
%  \streetaddress{Street2a Address2a}
%  \city{City2a}
%  \state{State2a}
%  \postcode{Post-Code2a}
  \country{USA}                   %% \country is recommended
}
%

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  We propose a runtime-assisted approach to enforce convergence in
  distributed executions of replicated data types. The key
  distinguishing aspect of our approach is that it guarantees
  convergence \emph{unconditionally} -- without requiring data type
  operations to satisfy algebraic laws such as commutativity and
  idempotence. Consequently, programmers are no longer obligated to
  prove convergence on a per-type basis. Moreover, our approach lets
  sequential data types be reused in a distributed setting by
  extending their implementations rather than refactoring them.  The
  novel component of our approach is a distributed runtime that
  orchestrates \emph{well-formed } executions that are guaranteed to
  converge. Despite the utilization of a runtime, our approach comes
  at no additional cost of latency and availability. Instead, we
  introduce a novel tradeoff against a metric called \emph{staleness},
  which roughly corresponds to the time taken for replicas to
  converge. We implement our approach in a system called \quark and
  conduct a thorough evaluation of its tradeoffs in the context of two
  case studies. 
%%We also briefly discuss the utility of our approach in
%%realizing causal coherency semantics in shared-memory concurrent
%%systems,
\end{abstract}

@@

%% Keywords
%% comma separated list
\keywords{Replication, MRDT, CRDT, Runtime, Convergence, Concurrent
Revisions, Causal Consistency, decentralized systems}  %% \keywords are mandatory in final camera-ready submission

@@

\bibliography{all}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
  We thank the anonymous reviewers and our shepherd Doug Lea for their
  thoughtful comments. First author has benefited from substantial
  discussions with Anmol Sahoo and Suresh Jagannathan. Anmol has also
  developed and maintained the OCaml-Scylla library, which is a key
  piece of infrastructure underlying \quark.
\end{acks}

%% Bibliography
%\bibliography{bibfile}

%% Appendix
% \appendix
% \section{Appendix}

>>>>

%\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{ifpdf}
\usepackage{graphicx}
%\usepackage[usenames,dvipsnames]{color}
\usepackage{stmaryrd}
%\usepackage[numbers]{natbib}
\usepackage{amsthm}
\usepackage{listings}          % format code
\usepackage{wrapfig}
\usepackage{textcomp}
\usepackage{tabularx}
\usepackage{color}
\usepackage{url}
\usepackage{tikz}
\usepackage{multirow,array}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage{balance} % To balance columns on last page

@@

%% Title information
\title[]{RunTime-Assisted Convergence in Replicated Data Types}
%% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

@@

%% Author with single affiliation.
\author{Gowtham Kaki}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{0000-0002-4189-3189}             %% \orcid is optional
\affiliation{
%  \position{Position1}
%  \department{Department1}              %% \department is recommended
  \institution{University of Colorado Boulder}            %% \institution is required
%  \streetaddress{Street1 Address1}
  \city{Boulder}
%  \state{State1}
%  \postcode{Post-Code1}
  \country{USA}                    %% \country is recommended
}
\email{gowtham.kaki@colorado.edu}          %% \email is recommended

%% Second Author
\author{Prasanth Prahladan}
%\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position2a}
%  \department{Department2a}             %% \department is recommended
  \institution{University of Colorado Boulder}           %% \institution is required
%  \streetaddress{Street2a Address2a}
  \city{Boulder}
%  \state{State2a}
%  \postcode{Post-Code2a}
  \country{USA}                   %% \country is recommended
}
\email{prasanth.prahladan@colorado.edu}         %% \email is recommended

%% Third Author
\author{Nicholas V. Lewchenko}
%\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position2a}
%  \department{Department2a}             %% \department is recommended
  \institution{University of Colorado Boulder}           %% \institution is required
%  \streetaddress{Street2a Address2a}
  \city{Boulder}
%  \state{State2a}
%  \postcode{Post-Code2a}
  \country{USA}                   %% \country is recommended
}
\email{nile1033@colorado.edu}         %% \email is recommended
%

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  We propose a runtime-assisted approach to enforce convergence in
  distributed executions of replicated data types. The key
  distinguishing aspect of our approach is that it guarantees
  convergence \emph{unconditionally} -- without requiring data type
  operations to satisfy algebraic laws such as commutativity and
  idempotence. Consequently, programmers are no longer obligated to
  prove convergence on a per-type basis. Moreover, our approach lets
  sequential data types be reused in a distributed setting by
  extending their implementations rather than refactoring them.  The
  novel component of our approach is a distributed runtime that
  orchestrates \emph{well-formed } executions that are guaranteed to
  converge. Despite the utilization of a runtime, our approach comes
  at no additional cost of latency and availability. Instead, we
  introduce a novel tradeoff against a metric called \emph{staleness},
  which roughly corresponds to the time taken for replicas to
  converge. We implement our approach in a system called \quark and
  conduct a thorough evaluation of its tradeoffs. 
%%We also briefly discuss the utility of our approach in
%%realizing causal coherency semantics in shared-memory concurrent
%%systems,
\end{abstract}

@@

%% Keywords
%% comma separated list
\keywords{Replication, MRDT, CRDT, Runtime, Convergence, Concurrent
Revisions, Causal Consistency, Decentralized Systems}  %% \keywords are mandatory in final camera-ready submission

@@

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
  We thank the anonymous reviewers and our shepherd Doug Lea for their
  thoughtful comments. First author has benefited from discussions
  with Anmol Sahoo and Suresh Jagannathan. Anmol has also developed
  and maintained the OCaml-Scylla library, which is a key piece of
  infrastructure underlying \quark.
\end{acks}

%% Appendix
\appendix
\section{Proofs}
\input{proofs}

%% Bibliography
\bibliography{all}

### 2022-04-25 related.tex
Committing edits

Mergeable Replicated Data Types (MRDTs) were introduced
in~\cite{mrdt}, where authors also demonstrate an approach for
deriving merge functions from first principles.  Despite resulting in
sensible merge semantics for several data types, their approach was
never shown to guarantee the convergence of the resultant MRDTs.
Indeed, as we demonstrate in Sec.~\ref{sec:motivation}, convergence
of unrestricted MRDT executions cannot be guaranteed due to the
presence of anomalies such as Fig.~\ref{fig:mrdt-exec-2} and
Fig.~\ref{fig:mrdt-exec-4}. \quark fixes this problem by extending
MRDTs with a runtime that limits their executions to those that are
guaranteed to converge.

@@

Several verification techniques, program analyses, and tools have been
proposed to reason about the program behavior in a weakly-consistent
distributed setting~\cite{bailis-vldb, alvaro-calm,
gotsman-popl16,redblue-atc, redblue-osdi, ecinec}. These techniques
treat replicated storage as a black box with a fixed pre-defined
consistency models. The focus is on assigning appropriate consistency
levels to operations so as to preserve application integrity. Such an
approach results in assigning sequential consistency (SC) to all
operations since the next weaker consistency model -- causal
consistency, is insufficient to guarantee convergence (as
Sec.~\ref{sec:motivation} demonstrates). Conversely, \quark cannot
reason about application-level invariants, such as $\C{balance}\ge 0$
in a banking application. Thus both approaches confer complimentary
benefits on application developers.

@@

Finally, the implementation of \quark bears resemblance to
Quelea~\cite{pldi15}. Both these systems are implemented as shim
layers on top of off-the-shelf distributed stores, and both offload
low-level concerns to the underlying store. Quelea however enforces
high-level invariants on applications built with CRDTs. Operations are
assumed to be commutative and applications are assumed to be
convergent. In contrast, \quark enforces convergence of data types
that are otherwise not guaranteed to converge.

>>>>

\balance
Mergeable Replicated Data Types (MRDTs) were introduced
in~\cite{mrdt}, where authors also demonstrate an approach for
deriving merge functions from first principles.  Despite resulting in
sensible merge semantics for several data types, their approach was
never shown to guarantee the convergence of the resultant MRDTs.
Indeed, as we demonstrate in Sec.~\ref{sec:motivation}, convergence
of unrestricted MRDT executions cannot be guaranteed due to the
presence of anomalies such as Fig.~\ref{fig:mrdt-exec-2} and
Fig.~\ref{fig:mrdt-exec-4}. \quark fixes this problem by extending
MRDTs with a runtime that limits their executions to those that are
guaranteed to converge.

@@

Several verification techniques, program analyses, and tools have been
proposed to reason about the program behavior in a weakly-consistent
distributed setting~\cite{bailis-vldb, alvaro-calm,
gotsman-popl16, pldi15, redblue-osdi}. These techniques
treat replicated storage as a black box with a fixed pre-defined
consistency models. The focus is on assigning appropriate consistency
levels to operations so as to preserve application integrity. Such an
approach results in assigning sequential consistency (SC) to all
operations since the next weaker consistency model -- causal
consistency, is insufficient to guarantee convergence (as
Sec.~\ref{sec:motivation} demonstrates). Conversely, \quark cannot
reason about application-level invariants, such as $\C{balance}\ge 0$
in a banking application. Thus both approaches confer complimentary
benefits on application developers.

@@

Finally, the implementation of \quark bears resemblance to
Quelea~\cite{pldi15} as both systems offload low-level concerns to an
underlying data store. While Quelea enforces high-level invariants on
applications built with CRDTs that are assumed to be convergent,
\quark enforces convergence of data types that are otherwise not
guaranteed to converge.
